CFF GCT Study

### 2025-09-23 Christine Chan/CFF GCT Study/cff_gct_analysis.qmd
figure updates for Monique
GCT formatting

---
title: "CFF GCT Study"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-height: 9
    fig-width: 12
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
library(rpart)
library(rpart.plot)
library(rms)
library(patchwork)
library(ggflowchart)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

df <- read.csv("./Christine Chan/CFF GCT Study/Data_Raw/CysticFibrosisScreen_DATA_LABELS_2024-04-19_1740.csv",

@@

  "White", "Black/African American", "American Indian/Alaska Native",
  "Asian Non-Indian", "Asian Indian", "Native Hawaiian/Other Pacific Islander"

@@

  df[, grep("Ethnic.Origin.Race", colnames(df))], 1,

@@

  df[, grep("Ethnicity", colnames(df))], 1,

@@

    "Weight (kg)" = Body.weight..kg., "Height (cm)" = Height..cm.,

@@

    "Second Highest FEV1pp" =
      Second.Highest.FEV1..12.months....FEV1.2nd.highest....,
    "Average of Highest FEV1pp" =
      Average.of.the.two.highest.FEV1.s...Average.FEV1.......,

@@

    "Type 2 diabetes in the family - Father" =
      factor(Type.2.diabetes.in.the.family...Father),
    "Type 2 diabetes in the family - Mother" =
      factor(Type.2.diabetes.in.the.family...Mother),

@@

    "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" =
      Number.of.other.relatives.with.diabetes..aunts..uncles..cousins..grandparents.

@@

    ID, HbA1c, Age, `Weight (kg)`, `Height (cm)`, BMI,
    Gender, Race, Ethnicity, `Pancreatic insufficient`,

@@

    `Highest FEV1pp`, `Second Highest FEV1pp`, `Average of Highest FEV1pp`,
    contains("GCT"), contains("OGTT "), "OGTT.Date.Time",

@@

df$`Height %ile` <- sds(df$`Height (cm)`,

@@

  item = "height2_20", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",

@@

df$`Weight %ile` <- sds(df$`Weight (kg)`,

@@

  item = "weight2_20", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",

@@

df$`BMI %ile` <- sds(df$BMI,

@@

  item = "bmi", ref = cdc.ref, type = "perc",
  male = "Male", female = "Female",

@@

df$Dysglycemia <- factor(df$Diagnosis,

@@

df$Diabetes <- factor(df$Diagnosis,

@@

df$CFRD <- factor(df$`OGTT 120 min` >= 200,

@@

  df[, c("OGTT 120 min", "GCT 60 min plasma")], 1, function(r) {

@@

df$`Got OGTT and GCT?` <- factor(df$`Got OGTT and GCT?`,

@@

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c + kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") + `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") + Gender + fe(Race) +
  fe(Ethnicity) + fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) + fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  `Highest FEV1pp` + `Second Highest FEV1pp` + `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(
    `Days From OGTT to GCT (Absolute Value)`,
    "N", "Nmiss2", "medianq1q3", "range"
  ) +
  fe(Diagnosis) + fe(Dysglycemia) + fe(Diabetes) + fe(CFRD)
t1a <- tableby(update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df, control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma and OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

@@

```{r results='asis'}
#| label: table1b
#| tbl-cap: Participant characteristics by CFRD status
#| message: false
t1b <- tableby(update(f, Diabetes ~ . - Diabetes),
  data = df, control = mycontrols
)
summary(t1b, pfootnote = T, digits = 2)
```

@@

```{r}
pairwise.t.test(df$`Average of Highest FEV1pp`, df$Diabetes)
pairwise.fisher.test(df$`Pancreatic insufficient`, df$Diabetes,
  p.adjust.method = "holm"
)
pairwise.t.test(df$HbA1c, df$Diabetes)
pairwise.wilcox.test(df$`GCT 60 min plasma`, df$Diabetes)
pairwise.wilcox.test(df$`GCT 60 min capillary`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 0 min`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 60 min`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 120 min`, df$Diabetes)
pairwise.wilcox.test(df$`Days From OGTT to GCT`, df$Diabetes)
```

@@

```{r results='asis'}
#| label: table1c
#| tbl-cap: Participant characteristics by dysglycemia
#| message: false
t1c <- tableby(update(f, Dysglycemia ~ . - Dysglycemia),
  data = df, control = mycontrols
)
summary(t1c, pfootnote = T, digits = 2)
```

@@

  pivot_longer(c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
    names_to = "OGTT Timepoint", values_to = "Glucose"

@@

plot_df$`OGTT Timepoint` <- factor(plot_df$`OGTT Timepoint`,

@@

ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",

@@

dotplot <- ggplot(df, aes(
  x = Diagnosis, y = `GCT 60 min plasma`,
  colour = Diagnosis
)) +

@@

  scale_color_manual(values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51"))

@@

ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot.png",

@@

  height = 4, width = 6, units = "in", dpi = 600,

@@

      `OGTT 0 min`, `OGTT 120 min`, `GCT 0 min plasma`, `GCT 60 min plasma`,
      `GCT 60 min capillary`

@@

    names_to = "OGTT Timepoint", values_to = "Glucose"

@@

    "OGTT 0 min", "OGTT 120 min", "GCT 0 min plasma", "GCT 60 min plasma",
    "GCT 60 min capillary"

@@

  labels = c("Fasting", "OGTT-2hG", "RPG", "GCTpl", "GCTcap")

@@

      `GCT 0 min plasma`, `GCT 60 min plasma`, `GCT 0 min capillary`,
      `GCT 60 min capillary`, HbA1c

@@

    names_to = "Measure", values_to = "Value"

@@

    axis.title.x = element_blank(), axis.text.x = element_blank(),

@@

    axis.title.y = element_blank(), axis.text.y = element_blank(),
    axis.title.x = element_blank(), axis.text.x = element_blank(),

@@

    legend.position = "none", plot.title = element_text(hjust = 0.5),
    axis.title.y = element_blank(), axis.text.y = element_blank(),
    axis.title.x = element_blank(), axis.text.x = element_blank(),

@@

    axis.title.y = element_blank(), axis.text.y = element_blank(), 

@@

    legend.title = element_blank(), plot.title = element_text(hjust = 0.5),

@@

    axis.title.x = element_blank(), axis.ticks.x = element_blank()

@@

dot_p <- dot_rpg + dot_rcg + dot_gctpl + dot_gctcap + dot_hba1c + 
  plot_layout(ncol = 5,guides = 'collect')
ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",

@@

  height = 9, width = 12, units = "in", dpi = 600,

@@

  stat_poly_eq(use_label(c("eq", "R2")),

@@

    label.y = "bottom", label.x = "right"

@@

ggsave("./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",

@@

```{r}
#| label: tbl-roc-dysglycemia
#| tbl-cap: Accuracy of different screening tests to predict dysglycemia using ROC curve analyses
#| message: false
gctpl_roc <- roc(Dysglycemia ~ `GCT 60 min plasma`, data = df)
gctcap_roc <- roc(Dysglycemia ~ `GCT 60 min capillary`, data = df)
ogtt0_roc <- roc(Dysglycemia ~ `OGTT 0 min`, data = df)
ogtt1_roc <- roc(Dysglycemia ~ `OGTT 60 min`, data = df)
rpg_roc <- roc(Dysglycemia ~ `GCT 0 min plasma`, data = df)
rcg_roc <- roc(Dysglycemia ~ `GCT 0 min capillary`, data = df)
a1c_roc <- roc(Dysglycemia ~ HbA1c, data = df)
roc_table <-
  lapply(
    list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
    function(roc) {
      coords(roc, "best", ret = "all")
    }
  )
roc_table <- do.call(rbind, roc_table)
rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 0 min", "OGTT 1 hr", "RPG", "RCG", "A1c")
gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

@@

```{r}
gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(gctpl_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(gctcap_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(ogtt0_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(ogtt1_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(rpg_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(rcg_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(a1c_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_dysglycemia.csv"
)
```

@@

```{r}
#| label: tbl-roc-cutoffs
#| tbl-cap: Sensitivity and specificity (95% bootstrap confidence intervals) to measure the performance of the glucose challenge test plasma glucose (GCTpl) to predict dysglycemia at different thresholds
#| message: false
gctpl_140 <- as.data.frame(ci.coords(gctpl_roc,
  x = 140,
  ret = c("sensitivity", "specificity")
))
gctpl_150 <- as.data.frame(ci.coords(gctpl_roc,
  x = 150,
  ret = c("sensitivity", "specificity")
))
gctpl_160 <- as.data.frame(ci.coords(gctpl_roc,
  x = 160,
  ret = c("sensitivity", "specificity")
))
gctpl_table <- do.call(rbind, list(gctpl_140, gctpl_150, gctpl_160))
rownames(gctpl_table) <- c(140, 150, 160)
gt(gctpl_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

@@

```{r}
#| label: tbl-roc-cfrd
#| tbl-cap: Accuracy of different screening tests to predict CFRD
#| message: false
gctpl_roc <- roc(CFRD ~ `GCT 60 min plasma`, data = df)
gctcap_roc <- roc(CFRD ~ `GCT 60 min capillary`, data = df)
ogtt0_roc <- roc(CFRD ~ `OGTT 0 min`, data = df)
ogtt1_roc <- roc(CFRD ~ `OGTT 60 min`, data = df)
rpg_roc <- roc(CFRD ~ `GCT 0 min plasma`, data = df)
rcg_roc <- roc(CFRD ~ `GCT 0 min capillary`, data = df)
a1c_roc <- roc(CFRD ~ HbA1c, data = df)
roc_table <-
  lapply(
    list(gctpl_roc, gctcap_roc, ogtt0_roc, ogtt1_roc, rpg_roc, rcg_roc, a1c_roc),
    function(roc) {
      coords(roc, "best", ret = "all")
    }
  )
roc_table <- do.call(rbind, roc_table)
rownames(roc_table) <- c("GCTpl", "GCTcap", "OGTT 0 min", "OGTT 1 hr", "RPG", "RCG", "A1c")
gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

@@

```{r}
gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(gctpl_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_CFRD.csv"
)
```

@@

```{r}
gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(gctcap_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_CFRD.csv"
)
```

@@

```{r}
gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(coords(ogtt0_roc, "all", ret = "all"),
  row.names = F, na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_CFRD.csv"
)
```

>>>>

---
title: "CFF GCT Study"
author: "Tim Vigers"
date: "now"
date-format: "MMMM D, YYYY [at] hh:mm a"
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    fig-height: 9
    fig-width: 12
    page-layout: full
editor: source
---

```{r setup}
#| include: false
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggpmisc)
library(pROC)
library(rspiro)
library(childsds)
library(arsenal)
library(gt)
library(reporttools)
library(rpart)
library(rpart.plot)
library(rms)
library(patchwork)
library(ggflowchart)
home_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(
  Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

df <- read.csv(
  "./Christine Chan/CFF GCT Study/Data_Raw/CysticFibrosisScreen_DATA_LABELS_2024-04-19_1740.csv",

@@

  "White",
  "Black/African American",
  "American Indian/Alaska Native",
  "Asian Non-Indian",
  "Asian Indian",
  "Native Hawaiian/Other Pacific Islander"

@@

  df[, grep("Ethnic.Origin.Race", colnames(df))],
  1,

@@

  df[, grep("Ethnicity", colnames(df))],
  1,

@@

    "Weight (kg)" = Body.weight..kg.,
    "Height (cm)" = Height..cm.,

@@

    "Second Highest FEV1pp" = Second.Highest.FEV1..12.months....FEV1.2nd.highest....,
    "Average of Highest FEV1pp" = Average.of.the.two.highest.FEV1.s...Average.FEV1.......,

@@

    "Type 2 diabetes in the family - Father" = factor(
      Type.2.diabetes.in.the.family...Father
    "Type 2 diabetes in the family - Mother" = factor(
      Type.2.diabetes.in.the.family...Mother

@@

    "Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)" = Number.of.other.relatives.with.diabetes..aunts..uncles..cousins..grandparents.

@@

    ID,
    HbA1c,
    Age,
    `Weight (kg)`,
    `Height (cm)`,
    BMI,
    Gender,
    Race,
    Ethnicity,
    `Pancreatic insufficient`,

@@

    `Highest FEV1pp`,
    `Second Highest FEV1pp`,
    `Average of Highest FEV1pp`,
    contains("GCT"),
    contains("OGTT "),
    "OGTT.Date.Time",

@@

df$`Height %ile` <- sds(
  df$`Height (cm)`,

@@

  item = "height2_20",
  ref = cdc.ref,
  type = "perc",
  male = "Male",
  female = "Female",

@@

df$`Weight %ile` <- sds(
  df$`Weight (kg)`,

@@

  item = "weight2_20",
  ref = cdc.ref,
  type = "perc",
  male = "Male",
  female = "Female",

@@

df$`BMI %ile` <- sds(
  df$BMI,

@@

  item = "bmi",
  ref = cdc.ref,
  type = "perc",
  male = "Male",
  female = "Female",

@@

df$Dysglycemia <- factor(
  df$Diagnosis,

@@

df$Diabetes <- factor(
  df$Diagnosis,

@@

df$CFRD <- factor(
  df$`OGTT 120 min` >= 200,

@@

  df[, c("OGTT 120 min", "GCT 60 min plasma")],
  1,

@@

df$`Got OGTT and GCT?` <- factor(
  df$`Got OGTT and GCT?`,

@@

```{r results='asis'}
#| label: table1a
#| tbl-cap: Participant characteristics by OGTT status
#| message: false
# Basic table formula
f <- ~ HbA1c +
  kwt(Age, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight (kg)`, "N", "Nmiss2", "medianq1q3", "range") +
  `Height (cm)` +
  kwt(BMI, "N", "Nmiss2", "medianq1q3", "range") +
  Gender +
  fe(Race) +
  fe(Ethnicity) +
  fe(`Pancreatic insufficient`) +
  fe(`Type 2 diabetes in the family - Father`) +
  fe(`Type 2 diabetes in the family - Mother`) +
  fe(`Brothers (number with diabetes)`) +
  fe(`Sisters (number with diabetes)`) +
  kwt(
    `Number of other relatives with diabetes (aunts, uncles, cousins, grandparents)`,
    "N",
    "Nmiss2",
    "medianq1q3",
    "range"
  ) +
  `Highest FEV1pp` +
  `Second Highest FEV1pp` +
  `Average of Highest FEV1pp` +
  kwt(`GCT 0 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 0 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min plasma`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`GCT 60 min capillary`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 0 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 60 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`OGTT 120 min`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Height %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Weight %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI %ile`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`BMI Value (Adults Only)`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(`Days From OGTT to GCT`, "N", "Nmiss2", "medianq1q3", "range") +
  kwt(
    `Days From OGTT to GCT (Absolute Value)`,
    "N",
    "Nmiss2",
    "medianq1q3",
    "range"
  ) +
  fe(Diagnosis) +
  fe(Dysglycemia) +
  fe(Diabetes) +
  fe(CFRD)
t1a <- tableby(
  update(f, `Got OGTT and GCT?` ~ . - `Got OGTT and GCT?`),
  data = df,
  control = mycontrols
)
summary(t1a, pfootnote = T, digits = 2)
# Limit to 189 after this
df <- df %>%
  filter(`Got OGTT and GCT?` == "Have GCT 1 Hour Plasma and OGTT 2 Hour") %>%
  select(-`Got OGTT and GCT?`)
```

@@

```{r results='asis'}
#| label: table1b
#| tbl-cap: Participant characteristics by CFRD status
#| message: false
t1b <- tableby(
  update(f, Diabetes ~ . - Diabetes),
  data = df,
  control = mycontrols
)
summary(t1b, pfootnote = T, digits = 2)
```

@@

```{r}
pairwise.t.test(df$`Average of Highest FEV1pp`, df$Diabetes)
pairwise.fisher.test(
  df$`Pancreatic insufficient`,
  df$Diabetes,
  p.adjust.method = "holm"
)
pairwise.t.test(df$HbA1c, df$Diabetes)
pairwise.wilcox.test(df$`GCT 60 min plasma`, df$Diabetes)
pairwise.wilcox.test(df$`GCT 60 min capillary`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 0 min`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 60 min`, df$Diabetes)
pairwise.wilcox.test(df$`OGTT 120 min`, df$Diabetes)
pairwise.wilcox.test(df$`Days From OGTT to GCT`, df$Diabetes)
```

@@

```{r results='asis'}
#| label: table1c
#| tbl-cap: Participant characteristics by dysglycemia
#| message: false
t1c <- tableby(
  update(f, Dysglycemia ~ . - Dysglycemia),
  data = df,
  control = mycontrols
)
summary(t1c, pfootnote = T, digits = 2)
```

@@

    c(`OGTT 0 min`, `OGTT 60 min`, `OGTT 120 min`),
    names_to = "OGTT Timepoint",
    values_to = "Glucose"

@@

plot_df$`OGTT Timepoint` <- factor(
  plot_df$`OGTT Timepoint`,

@@

ggsave(
  "./Christine Chan/CFF GCT Study/Dissemination/Figures/boxplot.png",

@@

dotplot <- ggplot(
  df,
  aes(
    x = Diagnosis,
    y = `GCT 60 min plasma`
  geom_jitter(width = 0.2) +

@@

  scale_color_manual(
    values = c("#264653", "#2a9d8f", "#e9c46a", "#f4a261", "#e76f51")

@@

ggsave(
  "./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot.png",

@@

  height = 4,
  width = 6,
  units = "in",
  dpi = 600,

@@

      `OGTT 0 min`,
      `OGTT 120 min`,
      `GCT 60 min plasma`

@@

    names_to = "OGTT Timepoint",
    values_to = "Glucose"

@@

    "OGTT 0 min",
    "OGTT 120 min",
    "GCT 60 min plasma"

@@

  labels = c("Fasting", "OGTT-2hG", "GCTpl")

@@

      `GCT 0 min plasma`,
      `GCT 60 min plasma`,
      `GCT 0 min capillary`,
      `GCT 60 min capillary`,
      HbA1c

@@

    names_to = "Measure",
    values_to = "Value"

@@

    axis.title.x = element_blank(),

@@

    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),

@@

    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.title.x = element_blank(),

@@

    axis.title.y = element_blank(),
    axis.text.y = element_blank(),

@@

    legend.title = element_blank(),

@@

    axis.title.x = element_blank(),

@@

dot_p <- dot_rpg +
  dot_rcg +
  dot_gctpl +
  dot_gctcap +
  dot_hba1c +
  plot_layout(ncol = 5, guides = "collect")
ggsave(
  "./Christine Chan/CFF GCT Study/Dissemination/Figures/dotplot2.png",

@@

  height = 9,
  width = 12,
  units = "in",
  dpi = 600,

@@

  stat_poly_eq(
    use_label(c("eq", "R2")),

@@

    label.y = "bottom",
    label.x = "right"

@@

ggsave(
  "./Christine Chan/CFF GCT Study/Dissemination/Figures/scatterplot.png",

@@

```{r}
#| label: tbl-roc-dysglycemia
#| tbl-cap: Accuracy of different screening tests to predict dysglycemia using ROC curve analyses
#| message: false
gctpl_roc <- roc(Dysglycemia ~ `GCT 60 min plasma`, data = df)
gctcap_roc <- roc(Dysglycemia ~ `GCT 60 min capillary`, data = df)
ogtt0_roc <- roc(Dysglycemia ~ `OGTT 0 min`, data = df)
ogtt1_roc <- roc(Dysglycemia ~ `OGTT 60 min`, data = df)
rpg_roc <- roc(Dysglycemia ~ `GCT 0 min plasma`, data = df)
rcg_roc <- roc(Dysglycemia ~ `GCT 0 min capillary`, data = df)
a1c_roc <- roc(Dysglycemia ~ HbA1c, data = df)
roc_table <-
  lapply(
    list(
      gctpl_roc,
      gctcap_roc,
      ogtt0_roc,
      ogtt1_roc,
      rpg_roc,
      rcg_roc,
      a1c_roc
    ),
    function(roc) {
      coords(roc, "best", ret = "all")
    }
  )
roc_table <- do.call(rbind, roc_table)
rownames(roc_table) <- c(
  "GCTpl",
  "GCTcap",
  "OGTT 0 min",
  "OGTT 1 hr",
  "RPG",
  "RCG",
  "A1c"
)
gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

@@

```{r}
gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(gctpl_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(gctcap_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(ogtt0_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(ogtt1_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(ogtt1_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_1_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(rpg_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(rpg_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RPG_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(rcg_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(rcg_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/RCG_dysglycemia.csv"
)
```

@@

```{r}
gt(data.frame(coords(a1c_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(a1c_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/A1c_dysglycemia.csv"
)
```

@@

```{r}
#| label: tbl-roc-cutoffs
#| tbl-cap: Sensitivity and specificity (95% bootstrap confidence intervals) to measure the performance of the glucose challenge test plasma glucose (GCTpl) to predict dysglycemia at different thresholds
#| message: false
gctpl_140 <- as.data.frame(ci.coords(
  gctpl_roc,
  x = 140,
  ret = c("sensitivity", "specificity")
))
gctpl_150 <- as.data.frame(ci.coords(
  gctpl_roc,
  x = 150,
  ret = c("sensitivity", "specificity")
))
gctpl_160 <- as.data.frame(ci.coords(
  gctpl_roc,
  x = 160,
  ret = c("sensitivity", "specificity")
))
gctpl_table <- do.call(rbind, list(gctpl_140, gctpl_150, gctpl_160))
rownames(gctpl_table) <- c(140, 150, 160)
gt(gctpl_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

@@

```{r}
#| label: tbl-roc-cfrd
#| tbl-cap: Accuracy of different screening tests to predict CFRD
#| message: false
gctpl_roc <- roc(CFRD ~ `GCT 60 min plasma`, data = df)
gctcap_roc <- roc(CFRD ~ `GCT 60 min capillary`, data = df)
ogtt0_roc <- roc(CFRD ~ `OGTT 0 min`, data = df)
ogtt1_roc <- roc(CFRD ~ `OGTT 60 min`, data = df)
rpg_roc <- roc(CFRD ~ `GCT 0 min plasma`, data = df)
rcg_roc <- roc(CFRD ~ `GCT 0 min capillary`, data = df)
a1c_roc <- roc(CFRD ~ HbA1c, data = df)
roc_table <-
  lapply(
    list(
      gctpl_roc,
      gctcap_roc,
      ogtt0_roc,
      ogtt1_roc,
      rpg_roc,
      rcg_roc,
      a1c_roc
    ),
    function(roc) {
      coords(roc, "best", ret = "all")
    }
  )
roc_table <- do.call(rbind, roc_table)
rownames(roc_table) <- c(
  "GCTpl",
  "GCTcap",
  "OGTT 0 min",
  "OGTT 1 hr",
  "RPG",
  "RCG",
  "A1c"
)
gt(roc_table, rownames_to_stub = T) %>% fmt_number(decimals = 3)
```

@@

```{r}
gt(data.frame(coords(gctpl_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(gctpl_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTpl_CFRD.csv"
)
```

@@

```{r}
gt(data.frame(coords(gctcap_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(gctcap_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/GCTcap_CFRD.csv"
)
```

@@

```{r}
gt(data.frame(coords(ogtt0_roc, "all", ret = "all"))) %>%
  fmt_number(n_sigfig = 4)
write.csv(
  coords(ogtt0_roc, "all", ret = "all"),
  row.names = F,
  na = "",
  file = "./Christine Chan/CFF GCT Study/Reports/Tables/OGTT_0_CFRD.csv"
)
```
